---
title: "Pivoting and summarizing data for presentation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pivoting and summarizing data for presentation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pknca.reporting)
```

# Pivoting Data

Sometimes presentation of data is easier when it is wide.  That may be done with
the `pivot_wider()` function for a `PKNCAconc` object.  Pivoting the data will,
by default, pivot to one group (i.e. subject and the other grouping factors) per
row and one column per nominal time point.  It will not work without the nominal
times given with the `time.nominal` argument to the PKNCAconc object.

After the data are pivoted, they may be summarized.  Summaries work on each
group across subjects.

# Example

```{r}
# Setup the data, in this case the nominal time is manually assigned by
# inspection of the raw data.
d_nom <-
  as.data.frame(datasets::Theoph) %>%
  mutate(
    time_nom =
      case_when(
        Time == 0~0,
        Time < 0.49~0.25,
        Time < 0.9~0.5,
        Time < 1.5~1,
        Time < 3~2,
        Time < 4.5~3.5,
        Time < 6~5,
        Time < 8~7,
        Time < 10~9,
        Time < 20~12,
        Time < 25~24,
        TRUE~NA_real_
      )
  )

# Make a PKNCAconc object (just like normal while ensuring that the time.nominal
# argument has been given)
o_conc <- PKNCAconc(d_nom, formula = conc~Time|Subject, time.nominal = "time_nom")
# Pivot the data wider
d_conc_pivot <- pivot_wider.PKNCAconc(o_conc)
knitr::kable(d_conc_pivot)

# Summarize that pivoted data
d_conc_pivot_summary <- summary(d_conc_pivot)
knitr::kable(d_conc_pivot_summary)
```
